---
title: "Figure 1"
output: html_notebook
---


```{r Load packages}
library(ggplot2); packageVersion("ggplot2")
library(plotrix)
library(nlme)
library(RColorBrewer)
library(GAD)
library(tidyverse)
library(plotly)
library(reshape2)

```

### Load data if you have not run "Get_working_datasets"

```{r}
diversity_data <- read.csv("Diversity_data.csv", header=T)
diversity_tib <- as_tibble(diversity_data)

```


### Aggregate data by medium
Each combination of carbon sources (growth medium) is replicated 3 times. 

```{r}
meanRich_tib <- diversity_tib %>%
   group_by(medium, FC_number, C_number) %>%
   summarise(mean_rich = mean(D0),
             SE_rich = std.error(D0))

# create numeric variable for medium 
meanRich_tib$medium_id <- as.numeric(levels(meanRich_tib$medium))[meanRich_tib$medium]

```

### Diversity in single resources 
List of single resources:

- Glucose
- Fructose
- Xylose
- Mannose
- Cellobiose
- Maltose
- Sucrose
- Citrate
- Fumarate
- Galacturonate
- Mannitol
- Sorbitol
- Glycerol
- Hydroxyproline
- Cellulose
- Starch.

Single resources include simple sugars, disaccharides, organic acids, sugar alcoohols, a compound similar to an amino acid (coomonly released by roots) and polysaccharides typical of soils. Citrate, fumarate and hydroxyproline are considered *gluconeogenic* resources (entering the central metabolic pathway via TCA), while all the others are *glycolytic* (entering from the glycolysis).

**One-way ANOVA test on data for single resources.** 
H0: average richness is similar across different growth media. H1: average richness of at least one media is different from the others. Tests provided both with and without the GAD package.

```{r}
diversity_1C <- filter(diversity_tib, C_number == 1)  # subset dataset
  
# anova_one_way <- aov(D0~medium, data = diversity_1C)
# summary(anova_one_way)
# 
# TukeyHSD(anova_one_way)

MEDIUM <- as.fixed(diversity_1C$medium)
ANOVA_model <- lm(D0 ~ MEDIUM, data = diversity_1C)
gad(ANOVA_model)

C.test(ANOVA_model)
snk.test(ANOVA_model, term = 'MEDIUM')

```

## Richness in single resources correlates with the number of metabolites predicted to be generated from them via rections inside the bacterial cells (generic ensemble of microbes). 
### Fig. 1 Panel B.

Metabolites are estimated from KEGG and MetaCyc databases (Akshit Goyal). We have one estimate for each growth medium (i.e. combination of carbon sources).

Plot relation between metabolite number and richness in single resources. Calculate Pearson Coefficient. 

```{r}
KEGG_data <- read.table("./Other_data/Kegg_data.txt", header=T)

KEGG_data_1C <- filter(KEGG_data, num_cs == 1)

meanRich_1C <- filter(meanRich_tib, C_number == 1)
#meanRich_1C$n_met <- KEGG_data_1C$met_number
#meanRich_1C$n_met_fba <- c(26,36,50,22,40,48,34,14,15,24,45,53,47,14,29,41)
n_met <- c(41,38,43,41,46,38,39,27,27,39,39,42,34,33,50,44)  # Curated number: need to fix this!

p_fig1_B <- ggplot(data=meanRich_1C, aes(n_met, y=mean_rich, fill=medium))+
  geom_errorbar(aes(ymin=mean_rich-SE_rich, ymax=mean_rich+SE_rich))+
  geom_point(size=5, shape=21)+
  theme_light()+
  geom_abline(aes(intercept=0, slope=0.62585), linetype="dashed", color="#990000")+
  #geom_smooth(method = "lm", linetype="dashed", color="#990000")+
  scale_fill_manual(values = colorRampPalette(brewer.pal(21, "Spectral"))(16))+
  theme(legend.position="none")+
  labs(y = "Observed richness", x="Estimated number of metabolites")+
  ylim(0,50)
p_fig1_B

pearson <- cor.test(n_met, meanRich_1C$mean_rich,
                    method="pearson")
pearson

```

```{r}
met_LM <- lm(mean_rich ~ 0 + n_met_fba, data=meanRich_1C)
summary(met_LM)
```

## How observed richness in constituent single resources compares to the observed richness in two-resource communities and predictions calculated as the union (sum without overlapping ASVs) or the maximum  of the richness in constituent singles. 
### Fig. 1 Panel C.
Having found large numbers of coexisting species in single resources, we expected that community diversity would increase rapidly if more resources were provided (union and max predictions).

Show that union and max predictions overestimate richness in 2-resource media: an example (glucose + hydroxyproline)

```{r}
dat_tmp_2C <- filter(meanRich_tib, C_number < 3 , medium == 1 | medium == 43 | medium == 56)
dat_tmp_2C <- dat_tmp_2C %>% select(!(medium_id))
dat_tmp_2C <-rows_insert(dat_tmp_2C, tibble(medium = "max", FC_number = "2", C_number = 2, mean_rich = 24.00000, SE_rich = 1.040833))
dat_tmp_2C <-rows_insert(dat_tmp_2C, tibble(medium = "union", FC_number = "2", C_number = 2, mean_rich = 29.88889, SE_rich = 0.9638529))


p_fig1_C <- ggplot(data=dat_tmp_2C, aes(x=FC_number, y=mean_rich, fill=medium))+
  geom_point(size=4, shape=21)+
  geom_errorbar(aes(x=C_number, y=mean_rich, ymin=mean_rich-SE_rich, ymax=mean_rich+SE_rich), width=.1)+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")+
  ylim(0,50)
p_fig1_C
```

## Observed average richness as a linear function of the number of supplied carbon sources
### Fig. 1 panel D
Predicted values based on union and maximum are obtained through a python script (name).

```{r}
predictions_df <- data.frame(union <- c(22.52083, 38.71, 63.48, 100.68, 149.36, 155.32), max = c(22.52083,28.61,33.99 ,38.85, 43.12, 43.53), C_number= c(1,2,4,8,15,16), FC_number = c("1","2","4","8","15","16"))

meanRich_Csource <- diversity_tib %>% 
  group_by(C_number) %>%
  summarise(mean_rich = mean(D0),
            SE_rich = std.error(D0))


p_fig1_D <- ggplot(data=predictions_df, aes(x=C_number, y=max))+
    geom_smooth(data=predictions_df, aes(x=C_number, y=union), method = lm, formula = y ~ splines::bs(x, 5), se = FALSE, color="black")+
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 5), se = FALSE, color="black")+
  geom_abline(slope=1, intercept =0, color="black")+
  geom_abline(slope=1.42583, intercept =20.65722, color="black")+
  geom_jitter(data=meanRich_tib, aes(x=C_number, y=mean_rich), size=1, shape=21, fill="gray", width=0.1)+
  geom_point(data=meanRich_Csource , aes(x=C_number, y=mean_rich), size=4, shape=21)+
  geom_errorbar(data=meanRich_Csource , aes(x=C_number, y=mean_rich, ymin=mean_rich-SE_rich, ymax=mean_rich+SE_rich))+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")+
  ylim(0,160)
p_fig1_D

```

### Extended data 
Get slopes and intercepts

```{r}
lm_rich <- lm(D0 ~ C_number, data=diversity_tib)
summary(lm_rich)


lm_rich_0.1 <- lm(D0_0.1 ~ C_number, data=diversity_tib)
summary(lm_rich_0.1)

lm_rich_fam <- lm(D0_family ~ C_number, data=diversity_tib)
summary(lm_rich_fam)


lm_rich_D1 <- lm(D1 ~ C_number, data=diversity_tib)
summary(lm_rich_D1)

lm_rich_D2 <- lm(D2 ~ C_number, data=diversity_tib)
summary(lm_rich_D2)
```

Plot

```{r}
plot_richnesses_df <- diversity_tib %>% 
  group_by(C_number) %>%
  summarise(mean_D1 = mean(D1),
            SE_D1 = std.error(D1),
            mean_D2 = mean(D2),
            SE_D2 = std.error(D2),
            mean_D0_0.1 = mean(D0_0.1),
            SE_D0_0.1 = std.error(D0_0.1),
            mean_D0_fam = mean(D0_family),
            SE_D0_fam = std.error(D0_family))

```


```{r}
library(gridExtra)

p_figS10A <- ggplot(data=diversity_tib, aes(x=C_number, y=D0_0.1))+
  geom_abline(slope=0.99952, intercept =11.59307, color="black")+
  geom_jitter(size=1, shape=21, fill="gray", width=0.1)+
  geom_point(data=plot_richnesses_df , aes(x=C_number, y=mean_D0_0.1), size=4, shape=21)+
  geom_errorbar(data=plot_richnesses_df , aes(x=C_number, y=mean_D0_0.1, ymin=mean_D0_0.1-SE_D0_0.1, ymax=mean_D0_0.1+SE_D0_0.1))+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")#+
  #ylim(0,32)
p_figS10A

p_figS10B <- ggplot(data=diversity_tib, aes(x=C_number, y=D0_family))+
  geom_abline(slope=0.42183, intercept =10.43812, color="black")+
  geom_jitter(size=1, shape=21, fill="gray", width=0.1)+
  geom_point(data=plot_richnesses_df , aes(x=C_number, y=mean_D0_fam), size=4, shape=21)+
  geom_errorbar(data=plot_richnesses_df , aes(x=C_number, y=mean_D0_fam, ymin=mean_D0_fam-SE_D0_fam, ymax=mean_D0_fam+SE_D0_fam))+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")#+
  #ylim(0,22)
p_figS10B

grid.arrange(p_figS10A, p_figS10B, nrow = 1)

p_figS10C <- ggplot(data=diversity_tib, aes(x=C_number, y=D1))+
  geom_abline(slope=0.31107, intercept =3.17260, color="black")+
  geom_jitter(size=1, shape=21, fill="gray", width=0.1)+
  geom_point(data=plot_richnesses_df , aes(x=C_number, y=mean_D1), size=4, shape=21)+
  geom_errorbar(data=plot_richnesses_df , aes(x=C_number, y=mean_D1, ymin=mean_D1-SE_D1, ymax=mean_D1+SE_D1))+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")

p_figS10D <- ggplot(data=diversity_tib, aes(x=C_number, y=D2))+
  geom_abline(slope=0.14559, intercept =2.33945, color="black")+
  geom_jitter(size=1, shape=21, fill="gray", width=0.1)+
  geom_point(data=plot_richnesses_df , aes(x=C_number, y=mean_D2), size=4, shape=21)+
  geom_errorbar(data=plot_richnesses_df , aes(x=C_number, y=mean_D2, ymin=mean_D2-SE_D2, ymax=mean_D2+SE_D2))+
  #geom_point(size=4, shape=22)+
  theme_bw()+
  theme(legend.position="none")

grid.arrange(p_figS10C, p_figS10D, nrow = 1)
```


# Multivariate regression

### Supplementary Fig. 18

```{r}
myDat <- diversity_tib

count_fun <- function(x){
  sum(x>0) 
}

myDat <- myDat %>%
   mutate(gluc_number = select(., citric_acid_c, fumaric_acid_c, proline_c) %>% apply(1, count_fun),
          glyc_number = select(.,glucose_c, fructose_c, xylose_c, mannose_c, cellobiose_c, maltose_c, sucrose_c, galacturonic_acid_c, mannitol_c, sorbitol_c, glycerol_c, cellulose_c, starch_c) %>% apply(1, count_fun))

diversity_glyc_gluc <- myDat %>%
   group_by(gluc_number,glyc_number,medium, FC_number, C_number) %>%
   summarise(mean_rich = mean(D0),
             SE_rich = std.error(D0))

LM.1 <- lm(mean_rich~gluc_number + glyc_number, data=diversity_glyc_gluc)
summary(LM.1)

LM.2 <- lm(mean_rich~C_number, data=diversity_glyc_gluc)
#summary(LM.2)

AIC(LM.1, LM.2)

```

### 3D plot

```{r}
my_df <- diversity_glyc_gluc
graph_reso <- 0.05

#Setup Axis
axis_x <- seq(min(my_df$glyc_number), max(my_df$glyc_number), by = graph_reso)
axis_y <- seq(min(my_df$gluc_number), max(my_df$gluc_number), by = graph_reso)

#Sample points
LM.1_surface <- expand.grid(glyc_number = axis_x,gluc_number = axis_y,KEEP.OUT.ATTRS = F)
LM.1_surface$pred_rich <- predict.lm(LM.1, newdata = LM.1_surface)
LM.1_surface <- acast(LM.1_surface, gluc_number ~ glyc_number, value.var = "pred_rich") #y ~ x

threeD_plot <- plot_ly(my_df, 
                     x = ~glyc_number, 
                     y = ~gluc_number, 
                     z = ~mean_rich,
                     text = ~C_number, 
                     type = "scatter3d", 
                     mode = "markers",
                     color= ~FC_number,
                     colors = c("#C2294A", "#FA9856", "#FEE08B", "#E6F598", "#94D4A4", "#4075B4"),
                     showscale = FALSE) 

threeD_plot <- threeD_plot %>% layout(
    title = "",
    scene = list(
      xaxis = list(title = "Number of glycolytic resources"),
      yaxis = list(title = "Number of gluconeogenic resources"),
      zaxis = list(title = "Richness")
    ))


threeD_PD_plot <- add_trace(p = threeD_plot,
                       z = LM.1_surface,
                       x = axis_x,
                       y = axis_y,
                       type = "surface", 
                       colorscale = "YlGnBu",
                       reversescale =T)

threeD_PD_plot
```


