---
title: "Figure 2"
output: html_notebook
---

## Distinguish between generalist and specialist ASVs

### Subset flat data at day 7 to obtain only single resource data and transform into presence-absence data for ASVs present in at least one resource.  
Probably not the best way to transform to presence-absence. This are utility codes to get datasets ready for the analysis on generalists and specialists. The main problem to solve is to exclude ASVs that are present in the dataset but absent from single resource communities. Finally count how many media each ASV is present in.

```{r}
df_ALL_RA_1C <- filter(df_ALL_RA, C_number == 1)

abundance_threshold = 0  #change to 0.001 to leave out low abund ASVs

ASV_SumAbund_1C <- df_ALL_RA_1C %>%
   group_by(OTU) %>%
   summarise(ASV_SumAbund = sum(Abundance,na.rm=TRUE))

ASV_SumAbund_1C_no0 <- filter(ASV_SumAbund_1C, ASV_SumAbund > abundance_threshold)
ASVtoKeep <- ASV_SumAbund_1C_no0$OTU

GenSpec_utility_df <- filter(df_ALL_RA_1C, OTU %in% ASVtoKeep)
GenSpec_utility_df$presence <- GenSpec_utility_df$Abundance
GenSpec_utility_df <- GenSpec_utility_df %>% mutate(presence = replace(presence, presence > 0, 1))

GenSpec_utility_df_2 <- GenSpec_utility_df %>%
   group_by(medium,OTU) %>%
   summarise(ASV_SumPresence = sum(presence,na.rm=TRUE))
GenSpec_utility_df_2 <- GenSpec_utility_df_2 %>% mutate(ASV_SumPresence = replace(ASV_SumPresence, ASV_SumPresence > 0, 1))

GenSpec_utility_df_3 <- GenSpec_utility_df_2 %>%
   group_by(OTU) %>%
   summarise(n_media = sum(ASV_SumPresence,na.rm=TRUE))   # count how many media (resource) each ASV is present in

p_fig2_A <- ggplot(GenSpec_utility_df_3 , aes(x=n_media))+
  geom_histogram(binwidth=1, fill="#CCCCCC", color="black")+
  scale_y_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))
   )
p_fig2_A
```

### Define generalist and specialist ASVs + Intermediates
This is simply based on how many media each is ASV is found

```{r}
Spec_df <- filter (GenSpec_utility_df_3, n_media <= 3 )
Spec_ASV <- unique(Spec_df$OTU)  # list of specialist ASVs

Gen_df <- filter (GenSpec_utility_df_3, n_media >= 13 )
Gen_ASV <- unique(Gen_df$OTU)  # list of generalist ASVs

Interm_df <- filter (GenSpec_utility_df_3, n_media > 3 & n_media < 13)
Interm_ASV <- unique(Interm_df $OTU)  # list of intermediate ASVs

```

### Find average number of generalists, specialists and intermediates in single and multiresource communities
Steps:

- Find number of **specialist**, **generalist** and **intermediate** ASVs in each replicate (distribution of resource strategies). 
- Compute average number of ASV for each strategy per medium (= mean over 3 reps). 
- Plot in singles and as a function of the number of C sources

1.
```{r}
### From 1 to 15 C sources
media <- c(1:74)
rep <- c("a","b","c")
n_media <- length(media)
n_rep <- length(rep)
dat <- df_ALL

# Specialists
out_spec_ASV=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_spec_ASV)<-c("r1","r2","r3")

for (i in 1:n_media){
  idx_1 <- media[i]
  g = dat[dat$medium==idx_1,]
    for (j in 1:n_rep){
  idx_2 <- rep[j]    
  m = g[g$rep==idx_2,]
  
       if (nrow(m)==0) {
        out_spec_ASV[i,j]= NaN
      } 
      else {
  
  ASV_no0 <- m[m$Abundance>0,]
  ASV_no0$OTU <- factor(ASV_no0$OTU)
  OTUs <- unique(ASV_no0$OTU)
  specialist_OTU <- length(Reduce(intersect, list(Spec_ASV,OTUs)))   #find elements in common

  out_spec_ASV[i,j] <- specialist_OTU
      }
  }
}

# Generalists
out_gen_ASV=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_gen_ASV)<-c("r1","r2","r3")

for (i in 1:n_media){
  idx_1 <- media[i]
  g = dat[dat$medium==idx_1,]
    for (j in 1:n_rep){
  idx_2 <- rep[j]    
  m = g[g$rep==idx_2,]
  
       if (nrow(m)==0) {
        out_gen_ASV[i,j]= NaN
      } 
      else {
  
  ASV_no0 <- m[m$Abundance>0,]
  ASV_no0$OTU <- factor(ASV_no0$OTU)
  OTUs <- unique(ASV_no0$OTU)
  generalist_OTU <- length(Reduce(intersect, list(Gen_ASV,OTUs)))   #find elements in common

  out_gen_ASV[i,j] <- generalist_OTU
      }
  }
}
  
# Intermediates
out_interm_ASV=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_interm_ASV)<-c("r1","r2","r3")

for (i in 1:n_media){
  idx_1 <- media[i]
  g = dat[dat$medium==idx_1,]
    for (j in 1:n_rep){
  idx_2 <- rep[j]    
  m = g[g$rep==idx_2,]
  
       if (nrow(m)==0) {
        out_interm_ASV[i,j]= NaN
      } 
      else {
  
  ASV_no0 <- m[m$Abundance>0,]
  ASV_no0$OTU <- factor(ASV_no0$OTU)
  OTUs <- unique(ASV_no0$OTU)
  intermediate_OTU <- length(Reduce(intersect, list(Interm_ASV,OTUs)))   #find elements in common

  out_interm_ASV[i,j] <- intermediate_OTU
      }
  }
}
  
### 16 C sources
media <- c(80:82)
rep <- c("a","b","c")
rep1 <- c("d","e","f")
rep2 <- c("g","h","i")
n_media <- length(media)
n_rep <- length(rep)
dat <- df_ALL_16C

# Specialists
out_spec_ASV_16=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_spec_ASV_16)<-c("r1","r2","r3")

  for (i in 1:n_media){
    g = dat[dat$medium==media[i],]
    for (j in 1:n_rep){
      m = g[g$rep %in% c(rep[j], rep1[j], rep2[j]),]
      ASV_no0 <- m[m$Abundance>0,]
      ASV_no0$OTU <- factor(ASV_no0$OTU)
      OTUs <- unique(ASV_no0$OTU)
      specialist_OTU <- length(Reduce(intersect, list(Spec_ASV,OTUs)))   #find elements in common
      
      out_spec_ASV_16[i,j] <- specialist_OTU
    }
  }  

# Generalists
out_gen_ASV_16=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_gen_ASV_16)<-c("r1","r2","r3")

  for (i in 1:n_media){
    g = dat[dat$medium==media[i],]
    for (j in 1:n_rep){
      m = g[g$rep %in% c(rep[j], rep1[j], rep2[j]),]
      ASV_no0 <- m[m$Abundance>0,]
      ASV_no0$OTU <- factor(ASV_no0$OTU)
      OTUs <- unique(ASV_no0$OTU)
      generalist_OTU <- length(Reduce(intersect, list(Gen_ASV,OTUs)))   #find elements in common
      
      out_gen_ASV_16[i,j] <- generalist_OTU
    }
  }  

# Intermediates
out_interm_ASV_16=as.data.frame(matrix(NA,nrow=n_media,ncol=n_rep))
  colnames(out_interm_ASV_16)<-c("r1","r2","r3")

  for (i in 1:n_media){
    g = dat[dat$medium==media[i],]
    for (j in 1:n_rep){
      m = g[g$rep %in% c(rep[j], rep1[j], rep2[j]),]
      ASV_no0 <- m[m$Abundance>0,]
      ASV_no0$OTU <- factor(ASV_no0$OTU)
      OTUs <- unique(ASV_no0$OTU)
      intermediate_OTU <- length(Reduce(intersect, list(Interm_ASV,OTUs)))   #find elements in common
      
      out_interm_ASV_16[i,j] <- intermediate_OTU
    }
  }

```

2.
```{r}
### Assemble datasets

#Spec  
Spec_ASV_Xrep_df <- as_tibble(rbind(out_spec_ASV,out_spec_ASV_16))
Spec_ASV_Xrep_df$medium <- c(1:77)
Spec_ASV_Xrep_df$strategy <- rep(c("Spec"),77)

Spec_ASV_Xrep_df <- Spec_ASV_Xrep_df %>% 
  pivot_longer(c(`r1`, `r2`, `r3`), names_to = "rep", values_to = "n")

Spec_ASV_Xmedium_df <- Spec_ASV_Xrep_df %>%
  group_by(strategy, medium) %>%
  summarise(mean_n = mean(n, na.rm=TRUE),
            SE_n = std.error(n, na.rm=TRUE))
Spec_ASV_Xmedium_df

# Gen
Gen_ASV_Xrep_df <- as_tibble(rbind(out_gen_ASV,out_gen_ASV_16))
Gen_ASV_Xrep_df$medium <- c(1:77)
Gen_ASV_Xrep_df$strategy <- rep(c("Gen"),77)

Gen_ASV_Xrep_df <- Gen_ASV_Xrep_df %>% 
  pivot_longer(c(`r1`, `r2`, `r3`), names_to = "rep", values_to = "n")

Gen_ASV_Xmedium_df <- Gen_ASV_Xrep_df %>%
  group_by(strategy, medium) %>%
  summarise(mean_n = mean(n, na.rm=TRUE),
            SE_n = std.error(n, na.rm=TRUE))
Gen_ASV_Xmedium_df

#Interm
Interm_ASV_Xrep_df <- as_tibble(rbind(out_interm_ASV,out_interm_ASV_16))
Interm_ASV_Xrep_df$medium <- c(1:77)
Interm_ASV_Xrep_df$strategy <- rep(c("Interm"),77)

Interm_ASV_Xrep_df <- Interm_ASV_Xrep_df %>% 
  pivot_longer(c(`r1`, `r2`, `r3`), names_to = "rep", values_to = "n")

Interm_ASV_Xmedium_df <- Interm_ASV_Xrep_df %>%
  group_by(strategy, medium) %>%
  summarise(mean_n = mean(n, na.rm=TRUE),
            SE_n = std.error(n, na.rm=TRUE))
Interm_ASV_Xmedium_df

```

3.
```{r}
### Order the richness data aggregated by medium_id (the numeric version of the factor medium) 
# I need this to retrieve the column "C_number" in the same order of the datasets with the strategies
meanRich_tib_ordered <- as_tibble(meanRich_tib[order(meanRich_tib$medium_id),])  

# patch all datasets together
strategies_Xmedium_df <- bind_rows(Gen_ASV_Xmedium_df, Interm_ASV_Xmedium_df, Spec_ASV_Xmedium_df)
strategies_Xmedium_df$C_number <- rep(c(meanRich_tib_ordered$C_number, 16, 16),3)

# single out single resources
strategies_Xmedium_1C_df <- filter(strategies_Xmedium_df, C_number == 1)
strategies_Xmedium_1C_df$dummy = rep(c("C_Generalist_ASVs", "B_Intermediate_ASVs", "A_Specialist_ASVs"), each= 16)
strategies_Xmedium_1C_df$tot_richness <- rep(c(meanRich_1C$mean_rich),3)
strategies_Xmedium_1C_df <- strategies_Xmedium_1C_df %>%
  arrange(strategy, tot_richness)
strategies_Xmedium_1C_df$asc_ordering <- rep(c(letters[1:16]), 3)

# compute error bar position
# strategies_Xmedium_1C_utility_df <- strategies_Xmedium_1C_df %>%
#   group_by(asc_ordering, medium) %>%
#   summarise(SE_n_pos = cumsum(mean_n))
# 
# strategies_Xmedium_1C_df <- strategies_Xmedium_1C_df %>%
#   arrange(dummy)
# strategies_Xmedium_1C_df$SE_n_pos <- strategies_Xmedium_1C_utility_df$SE_n_pos
# strategies_Xmedium_1C_df

p_fig2_B <- ggplot(data=strategies_Xmedium_1C_df, aes(asc_ordering, y=mean_n, fill=dummy))+
  geom_col(width=0.8, alpha=0.9)+
  #geom_errorbar(aes(ymin=SE_pos-SE, ymax=SE_pos+SE, group=strategy), width=0.2,  position = "identity")+
  scale_fill_manual(values = colorRampPalette(brewer.pal(4, "Spectral"))(3))+
  theme(legend.position="none")+
  labs(x = "Resource", y="Richness")
p_fig2_B

```

4.
```{r}
# aggregate per C number and strategy
strategies_XCnumber_df <- strategies_Xmedium_df %>%
                    group_by(strategy,C_number) %>% 
                        summarise(mean_strategies = mean(mean_n),
                                  SE_strategies = std.error(mean_n))

# sum over strategies and subtract from total richness to obtain the number of unseen ASVs in single resources
strategies_utility_df <- strategies_XCnumber_df %>% 
                         group_by(C_number) %>% 
                            summarise(tot_strategies_seen_in_single = sum(mean_strategies))

strategies_utility_df <- strategies_utility_df %>% 
  mutate(tot_richness = meanRich_Csource$mean_rich)

strategies_utility_df <- strategies_utility_df %>% 
  mutate(mean_strategies = tot_richness - tot_strategies_seen_in_single)
strategies_utility_df <- strategies_utility_df %>% 
  mutate(mean_strategies = if_else(mean_strategies < 0, 0, mean_strategies))
strategies_utility_df <- strategies_utility_df %>% 
  mutate(strategy = rep(c("Unseen_in_single"),6))
strategies_utility_df <- strategies_utility_df %>% 
  relocate(strategy, .before = C_number)

strategies_utility_df_1 <- strategies_XCnumber_df %>%
  select(strategy, C_number, mean_strategies)

strategies_utility_df_2 <- strategies_utility_df %>%
  select(strategy, C_number, mean_strategies)

strategies_plot_df <- bind_rows(strategies_utility_df_1, strategies_utility_df_2 )
strategies_plot_df

# add a column with strategy names in the order I want for the stacked bars plot
strategies_plot_df$dummy = rep(c("D_Generalist_ASVs", "C_Intermediate_ASVs", "B_Specialist_ASVs", "ASVs_Unseen_in_single"), each= 6)

p_fig2_G <- ggplot(data=strategies_plot_df, aes(C_number, y=mean_strategies, fill=dummy))+
  geom_col(width=0.8, alpha=0.9)+
  scale_fill_manual(values = colorRampPalette(brewer.pal(4, "Spectral"))(4))+
  theme(legend.position="none")+
  labs(x = "C_number", y="Richness")
p_fig2_G

```

## Look at the distribution of metabolites 
Compute the frequency of each metabolite across single resources

```{r}
metMap_1C_data <- read.table("./Other_data/Metabolites_1C.txt", header=T)

metMap_1C <- metMap_1C_data %>%
  pivot_longer(!medium, names_to = "metabolite", values_to = "presence")

met_count <-metMap_1C %>% 
    group_by(metabolite) %>%
    summarise(n_media = sum(presence))

p_fig2_C <- ggplot(met_count , aes(x=n_media))+
  geom_histogram(binwidth=1, fill="#CCCCCC", color="black")

p_fig2_C

```

### Define *core*, *intermediate* and *peripheral* metabolites
The thresholding is the same one we used for ASVs.

```{r}
core_met_tb <- met_count %>% filter(n_media >= 13)
core_met <- unique(core_met_tb$metabolite)

peripheral_met_tb <- met_count %>% filter(n_media <= 3)
peripheral_met <- unique(peripheral_met_tb$metabolite)

intermediate_met_tb <- met_count %>% filter(n_media > 3 & n_media < 13)
intermediate_met <- unique(intermediate_met_tb$metabolite)

```

### Find number of *core*, *intermediate* and *peripheral* metabolites generated from single-resource media
Steps:

- Find number of *core* (= common), *intermediate* and *peripheral* (= rare) metabolites in each medium 
- Plot in singles 

```{r}
media <- c(43:58)
n_media <- length(media)
dat <- metMap_1C

out_pheripheral_met <- integer(n_media)

for (i in 1:n_media){
  idx <- media[i]
  m = dat[dat$medium==idx,]
  
  met_no0 <- m[m$presence>0,]
  met_no0$metabolite <- factor(met_no0$metabolite)
  met <- unique(met_no0$metabolite)
  periph_met <- length(Reduce(intersect, list(peripheral_met,met)))   #find elements in common

  out_pheripheral_met[i] <- periph_met
}

out_core_met <- integer(n_media)

for (i in 1:n_media){
  idx <- media[i]
  m = dat[dat$medium==idx,]
  
  met_no0 <- m[m$presence>0,]
  met_no0$metabolite <- factor(met_no0$metabolite)
  met <- unique(met_no0$metabolite)
  cor_met <- length(Reduce(intersect, list(core_met,met)))   #find elements in common

  out_core_met[i] <- cor_met
}

out_intermediate_met <- integer(n_media)

for (i in 1:n_media){
  idx <- media[i]
  m = dat[dat$medium==idx,]
  
  met_no0 <- m[m$presence>0,]
  met_no0$metabolite <- factor(met_no0$metabolite)
  met <- unique(met_no0$metabolite)
  intermed_met <- length(Reduce(intersect, list(intermediate_met,met)))   #find elements in common

  out_intermediate_met[i] <- intermed_met
}

# Arrange in dataset
met_distribution_1C <- tibble(medium = rep(c(43:58),3), type = rep(c("a_Peripheral","b_Intermediate","c_Core"),each=16), number = c(out_pheripheral_met,out_intermediate_met,out_core_met), mean_rich = rep(c(meanRich_1C$mean_rich),3))
# letters in from of the name of the groups of metabolites are just for getting the right order in the plot

met_distribution_1C <- met_distribution_1C %>%
  arrange(type, mean_rich)
met_distribution_1C$asc_ordering <- rep(c(letters[1:16]), 3)

#Plot
p_fig2_D <- ggplot(data=met_distribution_1C, aes(asc_ordering, y=number, fill=type))+
  geom_col(width=0.8, alpha=0.9)+
  scale_fill_manual(values = colorRampPalette(brewer.pal(4, "Spectral"))(3))+
  theme(legend.position="none")+
  labs(x = "Resource", y="Number of metabolites")
p_fig2_D

```

### Plot preference score 
To verify the expectation that generalists and specialists also exhibit contrasting resource preferences in multi-resource environments, we calculated a resource-specificity score. For each ASV present in a single resource (target resource), the resource specificity score was calculated as the difference between the number of multi-resource media containing the target resource in which the ASV was found and the number of media not containing the target resource in which the ASV was found, divided by the total number of media in which the ASV was found. The score ranged from 1, indicating that the ASV was present only when the target resource was provided in a combination, to -1, implying that, although the ASV was found in the single resource, it was always absent when that resource was supplied with others. A score of 0 indicated that an ASV showed no specificity for that resource 

The data "Preference_score_ASVs.txt" was generated using "" (Hyunseok Lee).

```{r}
pref_score_ASV_df <- read.table("./Other_data/Preference_score_ASVs.txt", header=T)

pref_score_Xresource_df <- pref_score_ASV_df %>%
  group_by(resource, strategy) %>%
  summarise(mean_score = mean(preference),
            SE_score = std.error(preference))

pref_score_Xresource_df <-pref_score_Xresource_df %>%
  mutate(medium = case_when(resource=="glucose"~"43",
                            resource=="fructose"~"44",
                            resource=="xylose"~"45",
                            resource=="mannose"~"46",
                            resource=="cellobiose"~"47",
                            resource=="maltose"~"48",
                            resource=="sucrose"~"49",
                            resource=="citric_acid"~"50",
                            resource=="fumaric_acid"~"51",
                            resource=="galacturonic_acid"~"52",
                            resource=="mannitol"~"53",
                            resource=="sorbitol"~"54",
                            resource=="glycerol"~"55",
                            resource=="proline"~"56",
                            resource=="cellulose"~"57",
                            resource=="starch"~"58"
                            ))

pref_score_Xstrategy_df <- pref_score_Xresource_df %>%
  group_by(strategy) %>%
  summarise(avg_score = mean(mean_score),
            SE_score = std.error(mean_score))

p_fig2_F <- ggplot(data=pref_score_Xstrategy_df, aes(strategy, y=avg_score))+
  geom_jitter(data=pref_score_Xresource_df, size=2, shape=21, width= .1, aes(strategy, y=mean_score, fill=medium))+
  geom_col(alpha=0.5)+
  geom_errorbar(aes(ymin=avg_score - SE_score, ymax=avg_score + SE_score), width=0)+
  scale_fill_manual(values = colorRampPalette(brewer.pal(21, "Spectral"))(16)) +
  theme(legend.position="right")+
  labs(x = "", y="Mean resource-specificity score")
p_fig2_F

```

